### 2012-11-13 thesis/tex/2_background.tex
Final version?

%\begin{wrapfigure}{r}{0.38\textwidth}
%  \centering
%  \includegraphics[width=0.34\textwidth]{populationpyramide}
%  \caption{Age structure.}%Population pyramide of germany.}
%  \label{fig:populationpyramide}
%\end{wrapfigure}
In some countries, among New Zealand and Germany, the average age of population is constantly increasing. Owing to excellent achievements in medicine as well as higher living standards, humans can enjoy longer lifes. Also decreasing birth rates contribute to an upward drift of the dominating age range. As a result, the industrial sectors of professional health care and social institutions are required more than ever.
However, a strong deficit of professionals is denoted in the healthcare branch and availability is far behind demand. One reason for this might be below average low salaries for employees in the healthcare sector. But, just raising the salary will not lead to the desired results since the majority of the older people can not even afford a nurse to today's prices.

@@

The two newest robots IrobiQ and Cafero are shown in figure~\ref{fig:irobiq_cafero}.
These mobile 
\begin{wrapfigure}{r}{0.60\textwidth}
  \centering
  \includegraphics[width=0.56\textwidth]{irobiq_cafero}
  \caption{IrobiQ and Cafero, two robots of the healthcare project.}
  \label{fig:irobiq_cafero}
\end{wrapfigure}
robots are equipped with a touch display for presenting screens and receiving user input. Furthermore, they are equipped with additional external tools such as the blood pressure measurement device.
Applications running on the robots use the touch display to present screen dialogs to the user and use speech synthesis in order to give also accoustic feedback. User inputs or responses from devices cause dialogs to change. Figure~\ref{fig:screenflow_example} presents a small example of such a screen-flow based application, which uses response from a camera and user input for navigating through the screen-flow.

@@

For each state of the program, background actions can be defined. They are transparent to the user and run in background. Text messages can be sent or external devices can be acessed, for example. The latter is realized with web services~\cite{webservices} to obtain a modular design.
Some states can also have screen definitions with buttons, messages, videos, etc., on it. These are displayed to the user when the respective state is active. Once displayed, buttons can be clicked and trigger events.
The behaviour of events can be mapped in the expectedevents tag. It lists all events which are processable by the current state and defines the corresponding behaviour such as triggering transitions to other states.

@@

Robostudio is an editor fully written in Java and implemented as a NetBeans rich client application.
Several windows and views provide visual tools for editing and understanding robot behaviour as well as screen dialog design. As depicted in figure~\ref{fig:robostudio_vpe}, the editor integrates a \emph{State Navigator Window} (1) in the top left corner where all existing states are listed. Below an overview about all incoming and outgoing transitions is given. New states can be created and existing states can be deleted. A click on one particular state id in the \emph{State Navigator Window} will cause all other components to show all possible information regarding the selected state.
The \emph{Expected Events Editor Window} (2) in the bottom left can be used for specifying the expected events and their effects.
Heart and center of the editor is the \emph{UI Component Layout Editor Window} (3). It renders the screen dialog of the selected state and gives a preview of the screen presented on the robot during runtime. Components such as buttons, text boxes or images can be added, changed or removed via drag and drop.
They are provided by the \emph{UI Components Palette Window} (4) on the right side where the developer can choose from all supported UI components.
The \emph{Background Actions Editor Window} (5) lets the developer define background actions, which are executed transparently to the developer such as accessing external devices or sending messages.
Aiming to support the developer with helpful graphical tools, the \emph{State Transition Visualization Window} (6) gives an overview about all connections towards and from the selected state. It helps to get quickly an understanding of the local program behaviour.

@@

\section{Behaviour checking}
\label{sec:behaviourchecking}

The use and application of software is versatile and varies from text editing programs running on personal computers, over control software for high-precision applications in aeronautics, right up to trigger mechanisms for deathly weapons. For most of them safety properties are irrelevant, but a few are ranked highly dangerous to their environment and also to human lifes. In order to eliminate software malfunction there are different mathematical concepts for verifying software behaviour.
One of them is model checking which allows checking state machine programs for functional safety. Since the healthbot applications mentioned in section~\ref{sec:healthbotapplication} are realised wth state machines, this concept of behaviour checking is applicable and important for our kind of programs. At first the basic idea of state machines is described in section~\ref{sec:statemachines}, followed by an introduction into model checking in section~\ref{sec:modelcheckingandltl}.

@@

Each system status is represented by one \emph{state} which stands for a particular and unique constellation of system variables and properties. One state can be declared as \emph{start state} which is the very first state when the system starts working. The state representing the current system status is called \emph{current state}.
The current state changes whenever there is a modification to the system. Such transfers of the system from the current state to another one are defined by \emph{transitions}. Thus, they define the possible system behaviour.
All influences of the environment which cause a change to the system are called \emph{events}. They trigger transitions which transfer the system into a new state (current state). The fact that a state machine reacts to events is the reason why state machines are also called \emph{reactive systems}.

@@

	  \path[]
	  	(1) edge [bend right] node [left] {event b} (2)
	        edge [bend left] node [above] {event a} (3)
	    (2) edge [bend right] node [right] {event d} (3)
	    (3) edge [bend left] node [above] {event c} (1)
	        edge [loop right] node {event e} (3)
	    (4) edge [right] node [right] {} (1)
	    ;
	\end{tikzpicture}
  \caption{General example for a state machine.}
  \label{fig:statemachine}
\end{figure}
All states are drawn as circles which contain their names. Transitions are painted as arrows between source and destination states, and the regarding trigger event names are written next to it.

@@

The black dot defines the point of entry for this state machine, the door is closed but unlocked in the beginning. Through an event of the environment - in this case a human acting at the door - the door can be opened ,what makes ``opened'' the current state, and later be closed again. A turn of the key locks respectively blocks the door and no further opening or closing is possible. A second turn of the key un(b)locks the door again.

The short description given in this section does not contain a complete explanation about the concept of statemachines at all, but it gives a basic overview which is needed as background information for this work. Also the formal definition of state machines is not relevant and can be read in other literature~\cite{statemachines, statemachinesandformalmethods}.

@@

Whenever a software or hardware system has a special requirement in robustness and reliability, different formal methods are available to be used in order to verify the design.
Model checking is such a method which allows automatically verifying a system description against a specification by means of an exhaustive search of all states which could possibly be visited by the system during its execution. As a result, either the system description fulfills all propositions or a counter example is provided which gives evidence about a broken property. Figure~\ref{fig:modelchecking} gives a schematic view about the concept of model checking.

@@

For representing the system description - which can be a computer program, for example - a formal language is used. Also state machines are applicable to serve as a formal system description. In contrast, the specification used for model checking contains certain properties that express the desired and safe system behaviour in form of mathematical formulas. Linear temporal logic (LTL) is an example for a logic that is used to define such formulas on state machines. It is a temporal logic which can express formulas concerning the future of paths, for example, that something will eventually happen.

@@

\begin{description}
	\item[$\Box \varphi$] This formula expresses that $\varphi$ is true now and in all following states. The meaning of the rectangle is ``GLOBALLY''.
	\item[$\medcircle \varphi$] With a circle representing ``NEXT'', $\varphi$ will be true in the next state.
	\item[$\Diamond \varphi$] At least one time - now or eventually in a later state - $\varphi$ is or will be true. The diamond stands for ``FUTURE''.
\end{description}

@@

\begin{description}
	\item[$\varphi \mathcal{U} \psi$] The ``UNTIL'' operator makes $\varphi$ true at least until the first occurrence of a state where $\psi$ is true. Besides, $\psi$ will eventually be true.
	\item[$\varphi \mathcal{R} \psi$] It expresses that $\psi$ is true through the first occurrence of a state where $\varphi$ is true. Its meaning is $\varphi$ ``RELEASES'' $\psi$.
	\item[$\varphi \mathcal{W} \psi$] ``WEAK UNTIL'' represents the same as ``UNTIL'', except that an occurrence of $\psi$	is not necessary. In this case, $\varphi$ is true infinetely.
\end{description}

@@

The idea of a visual environment for the defining of system properties is not new and already pursued by some researchers. Although their focus does not completely match the subject of safety for healthcare robots, they present concepts of visualisation which might be relevant
% can be quite interesting 
for this work.
In section~\ref{sec:sisiruca}, the work of Sisirucá and Ionescu is presented. Section~\ref{sec:delbimbo} dexcribes a three dimensional concept by Del Bimbo et al., followed by an introduction of HomeTL in section~\ref{sec:hometl}.

@@

A first step towards a graphical tool for designing safety constraints is taken by Sisirucá and Ionescu~\cite{332301}. They developed an object-oriented graphical environment for visually creating temporal logic sentences and rules.
As shown in figure~\ref{fig:sisirucaandionescu}, the tool presents all operators as blocks which have inputs and outputs. 
\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.6\linewidth]{sisirucaandionescu} 
  \caption{The visual programming environment. An example showing the ruel construction process~\cite{332301}.}
  \label{fig:sisirucaandionescu}
\end{figure}
Whereas logical operators such as \emph{AND} and \emph{OR} have their boolean output and inputs represented by arrows pointing to the right, generic blocks for temporal expressions additionally have an input symbolized by an arrow pointing downward where the temporal operator itself is to be attached. By adding and connecting the operators a big expression graph can be built. In case that a diagram becomes too complex, an implode feature allows to compress the graph to a single box. It is also applicable for the reuse of diagrams.

The presented environment aims to be an effective tool for developing temporal logic sentences and rules. In fact, it gives a schematic overview of even complex formulas and allows substitution and reuse. The visual formalism is an alienation of LTL which might be suitable for researchers, but probably not for novices.

@@

Del Bimbo and his colleagues worked on a visual tool for temporal logic, too, but with a special focus on hierarchical representation of formulas~\cite{520786}.
It is based on computation tree logic (CTL) which is a branching-time logic with support for quantifiers. In contrast to LTL where properties are fulfilled by all possible paths, CTL expresses properties which depend on the future branching, for example, that at least one path is existend which has a particular property.
As indicated by figure~\ref{fig:delbimbo}, the visual tool provides all operators known from LTL 
%The visual tool provides visual operators as shown in figure~\ref{fig:delbimbo}.
\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.8\linewidth]{delbimbo} 
  \caption{Some of the visual operators of Del Bimbo et al.~\cite{520786}.}
  \label{fig:delbimbo}
\end{figure}
except that temporal operators are extended by an existential or universal quantifier (``at least on one path...'', ``on all paths...'').

Except propositions all operators are abstract and have placeholders representing their subformulas which can be connected to other operators.
%Here each node of a tree is described by an abstract operator whose subformulas are represented by placeholders which can be connected to other operators.
The positioning of these operators is organized in layers where the parent operator lays one level above its children. It can be compared to a tree upside down with the topmost operator as root.
This formula structure can be visualized as a 3D branching tree within a virual space. Figure~\ref{fig:delbimbo3d} shows the result of considerations done by Del Bimbo and his colleagues.

\begin{figure}[htbp]
  \centering
  \includegraphics[width=1\linewidth]{delbimbo3d} 
  \caption{Formula trees' 3D representation within a virual space~\cite{520786}.}
  \label{fig:delbimbo3d}
\end{figure}

The presented visual formalism directly maps the syntax of the temporal logic onto graphic symbols and composition patterns. The researchers found well expressive graphical representations for all operator especially the temporal ones, what makes the tool fairly intuitive. The tree dimensional visualisation gives a good overview of the structure, although it shuld not be used for editing. Computer monitors can naturally display only two dimensions and input devices such as mouse and keyboard are customised to it. Editing of the third dimension needs special gesture definitions and 3D modelling techniques the most users might not be familiar with. Thus, a 3D representation for editing should only be used if necessary. This perception has been tried to be considered in this work.

@@

The step towards a more abstract level of development is taken by HomeTL~\cite{4341725}, a visual formalism for the design of home based care. It allows healthcare professionals to specify rules and sequences of user actions in a quite nontechnical manner. Based on these conditions a monitoring system in a patient's home environment can detect and report abnormal situations within the daily routine.
The visual notation provides propositional variables and logical connectives, the latter is shown in figure~\ref{fig:hometloperators}. In contrast to LTL, the temporal operators of HomeTL use absolute time limits (seconds, for example) which are defined by $\tau$. Thus, not only the relative order of states is defined but also a time by which a particular action has to happen.

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.5\linewidth]{hometloperators} 
  \caption{Graphical representation of visual notation for temporal
operators within HomeTL user interface~\cite{4341725}.}
  \label{fig:hometloperators}
\end{figure}

This notation uses special meanings for the two dimensions: The horizontal dimension shows the temporal evolution of states, while the vertical dimension expresses the logical relationships at the respective current state. A programming environment allows editing and creating of constraints by using the atomar operators. Figure~\ref{fig:hometleditor} presents a snapshot of this editor.

>>>>

%\begin{wrapfigure}{r}{0.38\textwidth}
%  \centering
%  \includegraphics[width=0.34\textwidth]{populationpyramide}
%  \caption{Age structure.}%Population pyramide of germany.}
%  \label{fig:populationpyramide}
%\end{wrapfigure}
In some countries, among New Zealand and Germany, the average age of population is constantly increasing~\cite{altersstatistik,altersstatistik2}. Owing to excellent achievements in medicine as well as higher living standards, humans can enjoy longer lifes. Also decreasing birth rates contribute to an upward drift of the dominating age range. As a result, the industrial sectors of professional health care and social institutions are required more than ever.
However, a strong deficit of professionals is denoted in the healthcare branch and availability is far behind demand. One reason for this might be below average low salaries for employees in the healthcare sector. But, just raising the salary will not lead to the desired results since the majority of the older people can not even afford a nurse to today's prices.

@@

The two newest robots IrobiQ and Cafero are shown in figure~\ref{fig:irobiq_cafero}.
These mobile 
\begin{wrapfigure}{r}{0.60\textwidth}
  \centering
  \includegraphics[width=0.56\textwidth]{irobiq_cafero}
  \caption{IrobiQ and Cafero, two robots of the healthcare project.}
  \label{fig:irobiq_cafero}
\end{wrapfigure}
robots are equip\-ped with a touch display for presenting screens and receiving user input. Furthermore, they are equipped with additional external tools such as the blood pressure measurement device.
Applications running on the robots use the touch display to present screen dialogs to the user and use speech synthesis in order to give also accoustic feedback. User inputs or responses from devices cause dialogs to change. Figure~\ref{fig:screenflow_example} presents a small example of such a screen-flow based application, which uses response from a camera and user input for navigating through the screen-flow.

@@

For each state of the program, background actions can be defined. They are transparent to the user and run in background. Text messages can be sent or external devices can be accessed, for example. The latter is realized with web services~\cite{webservices} to obtain a modular design.
Some states can also have screen definitions with buttons, messages, videos, etc., on it. These are displayed to the user when the respective state is active. Once displayed, buttons can be clicked and trigger events.
The behaviour of events can be mapped in the expectedevents tag. It lists all events which are processable by the current state and defines the corresponding behaviour such as triggering transitions to other states.

@@

Robostudio is an editor fully written in Java and implemented as a NetBeans rich client application~\cite{netbeans}.
Several windows and views provide visual tools for editing and understanding robot behaviour as well as screen dialog design. As depicted in figure~\ref{fig:robostudio_vpe}, the editor integrates a \emph{State Navigator Window} (1) in the top left corner where all existing states are listed. Below an overview about all incoming and outgoing transitions is given. New states can be created and existing states can be deleted. A click on one particular state id in the \emph{State Navigator Window} will cause all other components to show all possible information regarding the selected state.
The \emph{Expected Events Editor Window} (2) in the bottom left can be used for specifying the expected events and their effects.
Heart and center of the editor is the \emph{UI Component Layout Editor Window} (3). It renders the screen dialog of the selected state and gives a preview of the screen presented on the robot during runtime. Components such as buttons, text boxes or images can be added, changed or removed via drag and drop.
They are provided by the \emph{UI Components Palette Window} (4) on the right side where the developer can choose from all supported UI components.
The \emph{Background Actions Editor Window} (5) lets the developer define background actions, which are executed transparently to the developer such as accessing external devices or sending messages.
Aiming to support the developer with helpful graphical tools, the \emph{State Transition Visualization Window} (6) gives an overview about all connections towards and from the selected state. It helps to get quickly an understanding of the local program behaviour.

@@

\section{Behaviour description and analysis}
\label{sec:behaviourchecking}

The use and application of software is versatile and varies from text editing programs running on personal computers, over control software for high-precision applications in aeronautics, right up to trigger mechanisms for deathly weapons. For most of them safety properties are irrelevant, but a few are ranked highly dangerous to their environment and also to human lifes. In order to eliminate software malfunction, there are different mathematical concepts for verifying software behaviour.
One of them is model checking which allows checking state machine programs for functional safety. Since the healthbot applications mentioned in section~\ref{sec:healthbotapplication} are realised with state machines, the concept of behaviour checking is applicable and important for this kind of programs. At first the basic idea of state machines is described in section~\ref{sec:statemachines}, followed by an introduction into model checking in section~\ref{sec:modelcheckingandltl}.

@@

Each system status is represented by one \emph{state} which stands for a particular and unique constellation of system variables and properties. One state can be declared as \emph{start state} which is the very first state when the system starts working. The state representing the current system status is called \emph{current state}.
The current state changes (maybe into itself) whenever there is a modification to the system. Such transfers of the system from the current state to another one are defined by \emph{transitions}. Thus, they define the possible system behaviour.
All influences of the environment which cause a change to the system are called \emph{events}. They trigger transitions which transfer the system into a new (current) state. The fact that a state machine reacts to events is the reason why state machines are also called \emph{reactive systems}.

@@

	  \path[]
	  	(1) edge [bend right] node [left] {event b} (2)
	        edge [bend left] node [above] {event a} (3)
	    (2) edge [bend right] node [right] {event d} (3)
	    (3) edge [bend left] node [above] {event c} (1)
	        edge [loop right] node {event e} (3)
	    (4) edge [right] node [right] {} (1)
	    ;
	\end{tikzpicture}
  \caption{General example for a state machine.}
  \label{fig:statemachine}
\end{figure}
All states are represented by circles which contain their names. Transitions are drawm as arrows between source and destination states, and the regarding trigger event names are written next to it.

@@

The black circle defines the point of entry for this state machine, the door is closed but unlocked in the beginning. By an event of the environment - in this case a human acting at the door - the door can be opened, which turns ``opened'' into the current state, and closed later on again. A turn of the key locks respectively blocks the door and no further opening or closing is possible. A second turn of the key unlocks respective unblocks the door again.

The short description given in this section does not contain a complete explanation about the concept of state machines at all, but it gives a basic overview which is needed as background information for this work.
Also the explanation of state machines above is no formal definition; it can be read found in literature~\cite{statemachines, statemachinesandformalmethods}.
%Also the formal definition of state machines is not relevant and can be found in literature~\cite{statemachines, statemachinesandformalmethods}.

@@

Whenever a software or hardware system has a special requirement regarding robustness and reliability, different formal methods are available to be used in order to verify the design.
Model checking~\cite{principles-of-model-checking} is such a method which allows automatically verifying a system description against a specification by means of an exhaustive search for all states which could possibly be visited by the system during its execution. As a result, either the system description fulfills all propositions or a counter example is provided which gives evidence about a broken property. Figure~\ref{fig:modelchecking} gives a schematic view about the concept of model checking.

@@

For representing the system description - which can be a computer program, for example - a formal language is used. Also state machines are applicable to serve as a formal system description. In contrast, the specification used for model checking contains certain properties that express the desired and safe system behaviour in form of mathematical formulas. Linear temporal logic (LTL)~\cite{principles-of-model-checking} is an example for a logic that is used to define such formulas on state machines. It is a temporal logic which can express formulas concerning the future of paths, for example, that something will eventually happen.

@@

\begin{description}
	\item[$\Box \varphi \textnormal{ }$] This formula expresses that $\varphi$ is true now and in all following states. The meaning of the rectangle is ``GLOBALLY''.
	\item[$\medcircle \varphi \textnormal{ }$] With a circle representing ``NEXT'', $\varphi$ will be true in the next state.
	\item[$\Diamond \varphi \textnormal{ }$] At least one time - now or eventually in a later state - $\varphi$ is or will be true. The diamond stands for ``FUTURE''.
\end{description}

@@

\begin{description}
	\item[$\varphi \mathcal{U} \psi$] The ``UNTIL'' operator makes $\varphi$ true at least until the first occurrence of a state where $\psi$ is true. Besides, $\psi$ will eventually be true.
	\item[$\varphi \mathcal{R} \psi$] It expresses that $\psi$ is true caused by the first occurrence of a state where $\varphi$ is true. Its meaning is $\varphi$ ``RELEASES'' $\psi$.
	\item[$\varphi \mathcal{W} \psi$] ``WEAK UNTIL'' represents the same as ``UNTIL'', except that an occurrence of $\psi$	is not necessary. In this case, $\varphi$ remains true infinitely.
\end{description}

@@

The idea of a visual environment for the defining of system properties is not new and already pursued by some researchers. Although their focus does not completely match the subject of safety for healthcare robots, they present concepts of visualization which might be relevant
% can be quite interesting 
for this work.
In section~\ref{sec:sisiruca}, the work of Sisirucá and Ionescu is presented. Section~\ref{sec:delbimbo} dexcribes a three dimensional concept by Del Bimbo et al., followed by an introduction of HomeTL in section~\ref{sec:hometl}.

@@

A first step towards a graphical tool for designing safety constraints is taken by Sisirucá and Ionescu~\cite{332301}. They developed an object-oriented graphical environment for creating temporal logic sentences and rules visually.
As shown in figure~\ref{fig:sisirucaandionescu}, the tool presents all operators as blocks which have inputs and outputs. 
\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.6\linewidth]{sisirucaandionescu} 
  \caption{The visual programming environment. An example showing the rule construction process~\cite{332301}.}
  \label{fig:sisirucaandionescu}
\end{figure}
Whereas logical operators such as \emph{AND} and \emph{OR} have each one boolean output and several boolean inputs represented by arrows pointing to the right, generic blocks for temporal expressions additionally have an input symbolized by an arrow pointing downward where the temporal operator itself is to be attached. By adding and connecting the operators a big expression graph can be built. In case that a diagram becomes too complex, an implode feature allows to compress the graph to a single box. It is also applicable for the reuse of diagrams.

The presented environment aims to be an effective tool for developing temporal logic sentences and rules. In fact, it gives a schematic overview of even complex formulas and allows substitution and reuse. The visual formalism is an alienation of LTL which might be suitable for researchers, but probably not for non-experts.

@@

Del Bimbo and his colleagues worked on a visual tool for temporal logic, too, but with a special focus on hierarchical representation of formulas~\cite{520786}.
It is based on computation tree logic (CTL) which is a branching-time logic with support for quantifiers. In contrast to LTL where properties are fulfilled by all possible paths, CTL expresses properties which depend on the future branching, for example, that at least one path exists which has a particular property.
As indicated by figure~\ref{fig:delbimbo}, the visual tool provides all operators known from LTL 
%The visual tool provides visual operators as shown in figure~\ref{fig:delbimbo}.
\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.8\linewidth]{delbimbo} 
  \caption{Visual representation of some of the operators of Del Bimbo et al.~\cite{520786}.}
  \label{fig:delbimbo}
\end{figure}
except that temporal operators are extended by an existential quantifier (e.g.\ ``at least on one path...'') or an universal quantifier (e.g.\ ``on all paths...'').

Except propositions all operators are abstract and have placeholders representing their subformulas which can be connected to other operators.
%Here each node of a tree is described by an abstract operator whose subformulas are represented by placeholders which can be connected to other operators.
The positioning of these operators is organized in layers where the parent operator lays one level above its children. It can be compared to a tree upside down with the topmost operator as root.
This formula structure can be visualized as a 3D branching tree within a virtual space. Figure~\ref{fig:delbimbo3d} shows the result of considerations done by Del Bimbo and his colleagues.

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.98\linewidth]{delbimbo3d} 
  \caption{Formula trees' 3D representation within a virtual space~\cite{520786}.}
  \label{fig:delbimbo3d}
\end{figure}

The presented visual formalism directly maps the syntax of the temporal logic onto graphic symbols and composition patterns. The researchers found well expressive graphical representations for all operator especially for the temporal ones, which makes the tool fairly intuitive. The three dimensional visualization gives a good overview of the structure, although it shuld not be used for editing. Computer monitors can naturally display only two dimensions and also input devices such as mouse and keyboard are customised to it. Editing of the third dimension needs special gesture definitions and 3D modelling techniques the most users might not be familiar with. Thus, a 3D representation for editing should only be used if necessary. This perception has been tried to be considered in this work.

@@

The step towards a more abstract level of development is taken by HomeTL~\cite{4341725}, a visual formalism for the design of home based care. It allows healthcare professionals to specify rules and sequences of actions in a quite nontechnical manner. Based on these conditions a monitoring system in a patient's home environment can detect and report abnormal situations within the daily routine.
The visual notation provides propositional variables and logical connectives, the latter is shown in figure~\ref{fig:hometloperators}. In contrast to LTL, the temporal operators of HomeTL use absolute time limits (seconds, for example) which are represented by $\tau$ in figure~\ref{fig:hometloperators}. Thus, not only the relative order of states is defined but also a time by which a particular action has to happen.

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.54\linewidth]{hometloperators} 
  \caption{Visual notation for temporal operators within HomeTL user interface~\cite{4341725}.}
  \label{fig:hometloperators}
\end{figure}

This notation gives special meanings to the two dimensions: The horizontal dimension shows the temporal evolution of states, while the vertical dimension (see~figure~\ref{fig:hometleditor}) expresses the logical relationships at the respective current state. A programming environment allows editing and creating of constraints by using the elementary operators shown in figure~\ref{fig:hometloperators}. Figure~\ref{fig:hometleditor} presents a snapshot of this editor.

### 2012-11-13 thesis/tex/3_goals.tex
Final version?

Based on the purpose of bringing the healthcare applications of the University of Auckland's healthcare project and certain safety standards closer together, and thereby to lay the foundation of aspired safety certifications, Robostudio should be extended with essential tools for this purpose. The visual programming environment is used for developing robotic applications in the form of state machines and shall provide a new interface in the future, which can be used by developers for evaluating safety properties of robot behaviour. On the one hand, this extension shall be seamlessly integrated in Robostudio. On the other hand, it should provide a clear API which allows easy integration also into other program environments. Thus, the new concepts are to be universal and not confined to Robostudio.

The goal is to introduce a visual language for defining safety constraints in state machine definitions of robot behaviour, that is suitable for people who are not experts in formal methods. The visual language should be easy to use and intuitive to read, in order to facilitate maintainability.
Hence, the graphical formalism should abstract from the usually complex and mathematical concept of conventional temporal logic but still be significantly expressive.

Furthermore some mechanism should be integrated aiming to support users in finding suitable constraints for a certain application. It is assumed that applications are modeled by state transition graphs. This modeling paradigm is used in many healthcare robots that employ dialogue systems for communicating with users~\cite{Bickmore2006}. Other possible fields might be shop assistant robots or entertainment robots. For such systems, the approach is suitable. It is based on heuristically analyzing state transition graphs and deriving constraints related to the graph structure. This work investigates a particular heuristic as a first step towards user support.

@@

\begin{itemize}
	\item The application will always eventually check the database for new reminder jobs. Thus it is ensured that pending reminders are processed.
%	\item Either a reminded medication is taken properly, or caregivers are notified that the patient refused medication intake.
	\item Whenever there is a pending reminder, medication will be finally taken or caregivers will get notified in case of the patient refusing medication intake.
% The next item is left out so that we can say, our heuristic found one we didn't think about yet :)
%	\item During the medication intake process the program can not switch to other services such as entertainment mode ore blood pressure measurement, for example.
\end{itemize}

@@

All the functionality mentioned above shall be combined in one editor and for evaluation purpose later be integrated into Robostudio, a programming environment for state machine editing.

>>>>

Based on the purpose of bringing the healthcare applications of the University of Auckland's healthcare project and certain safety standards closer together, and the\-re\-by laying the foundation of aspired safety certifications, Robostudio should be extended with essential tools for this purpose.

So the visual programming environment, used for developing robotic applications in the form of state machines, shall provide a new interface in the future, which can be used by developers for evaluating safety properties of robot behaviour. On the one hand, this extension shall be seamlessly integrated in Robostudio. On the other hand, it should provide a clear API which allows an easy integration also into other program environments. Thus, the new concepts are to be universal and not confined to Robostudio.

The main goal is to introduce such a visual language for defining safety constraints in state machine definitions of robot behaviour, that is suitable for people who are very familiar with the desired actions of the robot and the necessary safety conditions but are not experts in formal methods. The visual language should be easy to use and intuitive to read, in order to make the program easy to understand and to facilitate maintainability.
Hence, the graphical formalism should abstract from the usually complex and mathematical concept of conventional temporal logic but still be significantly expressive.

Furthermore some mechanism should be integrated aiming to support application developers in finding suitable constraints for a certain application. It is assumed that applications are modeled by state transition graphs. This modeling paradigm is used in many healthcare robots that employ dialogue systems for communicating with robot users~\cite{Bickmore2006}. Other possible fields might be shop assistant robots or entertainment robots. For such systems, the approach is suitable. It is based on heuristically analyzing state transition graphs and deriving constraints related to the graph structure. This work investigates a particular heuristic as a first step towards developer support.

@@

\begin{itemize}
	\item The application will always eventually check the database for new reminder jobs. Thus, it is ensured that pending reminders are processed.
%	\item Either a reminded medication is taken properly, or caregivers are notified that the patient refused medication intake.
	\item Whenever there is a pending reminder, either medication will be finally taken or caregivers will get notified in case of the patient refusing medication intake.
% The next item is left out so that we can say, our heuristic found one we didn't think about yet :)
%	\item During the medication intake process the program can not switch to other services such as entertainment mode ore blood pressure measurement, for example.
\end{itemize}

@@

All the functionality mentioned above shall be combined in one editor and subsequently for evaluation purpose be integrated into Robostudio, a programming environment for state machine editing.
